---
title: "p8105_hw1"
output: github_document
date: "2023-09-21"
---

# Problem 1
## part 0 - setup
call the library **moderndive** and **tidyverse** and install the dataset *early_january_weather* (library **tidyverse** is also used in problem 2)
```{r message = FALSE}
library(moderndive)
library(tidyverse)
data("early_january_weather")
problem1_df = early_january_weather
```

## part 1 - description of the dataset (partial code and dataset description)
```{r}
variable_name = c(names(problem1_df))
num_row = nrow(problem1_df)
num_col = ncol(problem1_df)
mean_temp = mean(pull(problem1_df, temp))

data_year = unique(pull(problem1_df, year))
data_site = unique(pull(problem1_df, origin))
start_day = with(problem1_df, min(day))
end_day = with(problem1_df, max(day))
```
The installed dataset contains variables *`r variable_name`*; it 
contains *`r num_row`* rows and *`r num_col`* columns. The
mean temperature in the reported days is *`r round(mean_temp, digits = 2)`*. 
**Additonal information**: The data is collected in the year *`r data_year`* in 
*`r data_site`* from day *`r start_day`* to *`r end_day`*.

## part 2 - make a scatterplot plot
```{r message = FALSE}
ggplot(problem1_df, aes(x = time_hour, y = temp, colour = humid)) + geom_point()

ggsave("problem1_scatterplot.pdf")
```
Describe the pattern of the plot: the temperature increases in general as time
goes, even though it fluctuates during the period of data collection.The humid
level also tends to increase (but also fluctuates).



# Problem 2
## part 1 - create a dataframe and try to take the mean of each variable
```{r}
problem2_df = tibble(
  norm_sample = rnorm(10),
  norm_samp_log = norm_sample > 0,
  vec_char = c("apple", "banana", "peach", "flash", "milk", "cookie", "sparkle",
               "tigger", "melon", "cauliflower"),
  fac_char = factor(c("dog", "dog", "dog", "sloth", "dog", "cat", "cat", "cat", 
                      "cat", "dog"))
)

mean(pull(problem2_df, norm_sample))
mean(pull(problem2_df, norm_samp_log))
mean(pull(problem2_df, vec_char))
mean(pull(problem2_df, fac_char))


```
When we try to take the mean of each variable, we can take the mean of the numeric
and the logical variable. We cannot take the mean of a character or a factor vector.

## part 2 - apply the as.numeric function to the logical, character, and factor variables
```{r results = 'hide'}
as.numeric(pull(problem2_df, norm_samp_log))
as.numeric(pull(problem2_df, vec_char))
as.numeric(pull(problem2_df, fac_char))
```
We can return numeric values for the logical and factor vector, but "NA" will
be returned for the character vector, and that might be the reason why we cannot
take the mean of a character variable. As for the factor vector, even though 
numeric values are returned after **as.numeric** function is returned, we could 
not calculate the mean, that might because the numeric values returned for 
the factor 'levels' do not have a 'meaning'.





